knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
##Predicting Quality of Exercise Using Personal Activity Devices
Through devices such as Nike FuelBand, Fitbit, and Jawbone Up, we can collect a large amount of data about personal activity. People regularly measure how much of a particular activity they do, but the information about how well they do it is missing.

Six participants were asked to make the barbell lifts correctly and incorrectly in 5 different ways. The goal of this project is to use data from accelerometers on the belt, forearm, arm, and dumbbell of these participants to predict how they did the exercise. The data for this project can be found here: http://groupware.les.inf.puc-rio.br/har
##Loading the data
Since we're only interested in the accelerometer data, let's load the data and only include the accelerometer data, the names of the participants, and the 5 ways in which they did the barbell lift exercises (classe). We'll also make sure there are no missing values.
library(tidyverse)
library(caret)

```{r}
library(tidyverse)
library(caret)

data <- read.csv(file.choose(),header = TRUE) ##choose pml-training .csv file from the File Explorer

data_new <- data %>% select(user_name, contains("accel"), classe) %>% select(!contains("var"))
```
# Explore the structure of the dataset
str(data_new)

# Check for missing values
sum(is.na(data_new))
#Using Cross-Validation
We will partition the training dataset into a training (75% of the original training set) and testing dataset to evaluate the model after we build it.
```{r}
set.seed(12345)

inTrain <- createDataPartition(data_new$classe, p = 0.75, list = FALSE)

training <- data_new[inTrain,]
testing <- data_new[-inTrain,]
dim(training) ; dim(testing) 
             ```{r}
Building the Model
We have at least 18 predictors which may not all be useful in our model. We can do a principal component analysis to further reduce our predictors.
library(caret)

# Preprocess data using PCA
princo <- preProcess(training, method = "pca", thresh = 0.8)

training_pca <- predict(princo, training)
summary(training_pca)
There are 6 principal components now, so let's build our model. I decided to use a Random Forest model for these data because of its flexibility and ability to capture complex relationships.
start.time <- Sys.time()
library(randomForest)

set.seed(12345)
mod_rf <- train(classe ~ ., method = "rf", data = training_pca)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken

mod_rf$finalModel
#Testing the Model
new_test <- predict(princo, testing)

pred_rf <- predict(mod_rf, new_test)

conmat_rf <- confusionMatrix(as.factor(testing$classe), pred_rf)

conmat_rf

